A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate

### 2020-05-06 Manuscript/Manuscript-5.Rmd
Supp INfo

Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa. This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. Other approaches for interpreting $\lambda$ are more statistical. For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019]. Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019]. Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature. For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait. Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.

@@

Second, while estimates of $\lambda$ are within a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data. For this to be the case, $\lambda$ must be a statistical effect size. Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988]. Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b]. Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994]. As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable. For the case of $\lambda$, this has not yet been explored.  \hfill\break

@@

*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions. Predictably, precision improved as the number of species increased (Figure 2). This confirmed earlier findings of Boettiger et al. [-@Boettiger_et_al2012], and adhered to parametric statistical theory. However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal. Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal. The worst precision was observed at intermediate levels of phylogenetic signal ($\lambda_{in}\approx0.5$), while precision improved as input levels approached the extremes of $\lambda$'s range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$). Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values. Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$). Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$). These findings were particularly unsettling when considered in light of our literature survey. Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1). This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these mid-range empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal). \hfill\break

@@

where $\theta_{obs}$ is the observed test statistic, $E(\theta)$ is its expected value under the null hypothesis, and $\sigma_\theta$ is its standard error [@Glass1976; @Cohen1988;@Rosenthal1994]. $Z_{\theta}$ expresses the magnitude of the effect in $\theta_{obs}$ by transforming the original test statistic to its standard normal deviate [@Glass1976; @Kelley2012]. Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory. However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures. \hfill\break

@@

where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $log(Kappa)$ obtained via permutation. Here the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018]. Similarly, an effect size based on $\lambda$ could be envisioned as:

@@

To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$). Results are found in Figure 4 (additional results are found in the Supporting Information). Here two things are evident. First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B). Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$. Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$. This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.  Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C). This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$. \hfill\break

@@

Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared to another. As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:

@@

To demonstrate the utility of $\hat{Z}_{12}$ we performed an analysis of the strength of phylogenetic signal in two phenotypic datasets from species of plethodontid salamander. The data were part of a series of studies examining macroevolutionary trends in phenotypic diversification in this group [@Baken2019; @Baken2020]. Our dataset contained surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) for 305 species (Figure 5A). For SA:V, 11 linear body measurements from 2,781 individuals were taken, from which estimates of the surface area and volume of the head, body, and tail were calculated and subsequently combined to arrive at the SA:V for each indiviual [for mathematical details see @Baken2020]. Species means were then obtained. Likewise, body size (SVL) and body width (BW) measurements were taken from 3,371 individuals, and species means of relative body width ($\frac{BW}{SVL}$) were calculated [data from @Baken2019]. A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the dataset, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species. The phylogenetic signal in each trait was then characterized using $Kappa$, which was converted to its effect size ($Z_K$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph]. Finally, the strength of phylogenetic signal was compared across traits using $\hat{Z}_{12}$ as described above (**to be incorporated in `geomorph` upon manuscript acceptance**). \hfill\break

*Results.* Both SA:V and relative body width displayed significant phylogenetic signal ($Kappa_{SA:V}=0.7608$; $P=0.001$; $Kappa_{BW/SVL}=0.2515$; $P=0.001$).  For both phenotypic traits, $K_{obs}$ differed markedly from their corresponding permutation distributions, which were found to overlap almost perfectly (Figure 5B). However, while both traits displayed significant phylogenetic signal, there was nearly a four-fold difference in the magnitude of their effect sizes, with SA:V displaying the greater phylogenetic signal (Figure 5C). Using the two-sample test statistic above, this difference was found to be highly significant ($\hat{Z}_{12}=4.13$; $P=0.000036$). Thus it may be concluded that SA:V displays significantly stronger phylogenetic signal than does relative body width, and that shared evolutionary history has strongly influenced trait covariation among taxa for SA:V. Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020]. Thus, because of this macroevolutionary association, strong phylogenetic signal in SA:V is to be expected.

@@

As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal. $Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean. We applied this concept to both $\lambda$ and $Kappa$, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data. First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal. Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomite changes in the values of $Z_K$. Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data. We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect. \hfill\break

Based on the effect size $Z_K$, we then proposed a two-sample test, which provides a statistical means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another. Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $Kappa$ [e.g.,  @Blomberg_et_al2003]. However, $Kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information]. Thus, $Kappa$ should not be considered a standardized effect size that measures the strength of phylogenetic signal on a common scale. By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets. Thus when viewed from this perspective, the approach developed here is statistically aligned with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons. As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.  \hfill\break

One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012]. Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation. This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$. For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$). Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a]. Likewise, as $Z_K$ is a standard deviation unit (which accounts for variation, sample size, and the phylogeny), tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal across datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence, and observed variation, are taken into account in the generation of the empirical sampling distribution via permutation.    \hfill\break

** Finally... Need Closing paragraph. **

>>>>

Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa. This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. Other approaches for interpreting $\lambda$ are more statistical. For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019]. Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019]. Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature. For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found from a literature survey we conducted in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait. Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.

@@

Second, while estimates of $\lambda$ are within a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data. For this to be the case, $\lambda$ must be a statistical effect size. Effect sizes are a measure of the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988]. Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b]. Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994]. As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable. For the case of $\lambda$, this has not yet been explored.  \hfill\break

@@

*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions. Predictably, precision improved as the number of species increased (Figure 2). This confirmed earlier findings of Boettiger et al. [-@Boettiger_et_al2012], and adhered to parametric statistical theory. However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal. Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal. The worst precision was observed at intermediate levels of phylogenetic signal ($\lambda_{in}\approx0.5$), while precision improved as input levels approached the extremes of $\lambda$'s range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$). Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values. Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$). Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$). These findings were particularly unsettling when considered in light of our literature survey. Over one quarter of the $\lambda$ estimates published in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1). This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these mid-range empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal). \hfill\break

@@

where $\theta_{obs}$ is the observed test statistic, $E(\theta)$ is its expected value under the null hypothesis, and $\sigma_\theta$ is its standard error [@Glass1976; @Cohen1988;@Rosenthal1994]. $Z_{\theta}$ expresses the magnitude of the effect in $\theta_{obs}$ by transforming the original test statistic to its standard normal deviate [@Glass1976; @Kelley2012]. Typically, $\theta_{obs}$ and $\sigma_\theta$ are estimated from the data, while $E(\theta)$ is obtained from the distribution of $\theta$ derived from parametric theory. However, recent advances in resampling theory [@Collyer_et_al2015a;@AdamsCollyer2016; @CollyerAdams2018;@AdamsCollyer2019b] have shown that $E(\theta)$ and $\sigma_\theta$ may also be obtained from an empirical sampling distribution of $\theta$ obtained from permutation procedures. \hfill\break

@@

where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $log(Kappa)$ obtained via permutation. Note that the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018]. \hfill\break

An effect size based on $\lambda$ could be envisioned, which is found as:

@@

To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$, Figure 4: additional results in the Supporting Information). Here two things are evident. First, estimates of $Z_K$ linearly track the input phylogenetic signal whereas estimates of $Z_\lambda$ do not (Figure 4A,B). Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$. Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$. This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.  Here coefficients of variation in the precision of $Z_K$ were up to an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C), implying that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$. \hfill\break

@@

Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait than another. As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:

@@

To demonstrate the utility of $\hat{Z}_{12}$ we quantified and compared the strength of phylogenetic signal of two phenotypic traits across species of plethodontid salamander. The data were part of a series of studies examining macroevolutionary trends in phenotypic diversification in this group [@Baken2019; @Baken2020]. Our dataset contained surface area to volume ratios (SA:V) and relative body width ($\frac{BW}{SVL}$) for 305 species (Figure 5A). For SA:V, 11 linear body measurements from 2,781 individuals were taken, from which estimates of the surface area and volume of the head, body, and tail were calculated and subsequently combined to arrive at the SA:V for each indiviual [for mathematical details see @Baken2020]. Species means were then obtained. Likewise, body size (SVL) and body width (BW) measurements were taken from 3,371 individuals, and species means of relative body width ($\frac{BW}{SVL}$) were calculated [data from @Baken2019]. A time-dated molecular phylogeny for the group [@Bonett2017] was then pruned to match the species in the dataset, resulting in a phylogeny and corresponding phenotypic dataset containing 305 species. The phylogenetic signal in each trait was then characterized using $Kappa$, which was converted to its effect size ($Z_K$) using `geomorph` 3.2.1 [@AdamsOtarola2013; @AdamsGeomorph]. Finally, the strength of phylogenetic signal was compared across traits using $\hat{Z}_{12}$ as described above (**to be incorporated in `geomorph` upon manuscript acceptance**). \hfill\break

*Results.* Both SA:V and relative body width displayed significant phylogenetic signal ($Kappa_{SA:V}=0.7608$; $P=0.001$; $Kappa_{BW/SVL}=0.2515$; $P=0.001$).  For both phenotypic traits, $K_{obs}$ differed markedly from their corresponding permutation distributions, which were found to overlap almost perfectly (Figure 5B). However, while both traits displayed significant phylogenetic signal, there was nearly a four-fold difference in the magnitude of their effect sizes, with SA:V displaying the greater phylogenetic signal (Figure 5C). Using the two-sample test statistic above, this difference was found to be highly significant ($\hat{Z}_{12}=4.13$; $P=0.000036$). Thus it may be concluded that SA:V displays significantly stronger phylogenetic signal than does relative body width, and that shared evolutionary history has strongly influenced trait covariation among taxa for SA:V. Biologically, this observation corresponds with the fact that tropical species -- which form a monophyletic group within plethodontids -- display greater variation in SA:V which covaries with disparity in their climatic niches [@Baken2020]. We hypothesize that because of this macroevolutionary association, strong phylogenetic signal in SA:V is observed.

@@

As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal. $Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean. We applied this concept to both $\lambda$ and $Kappa$, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data. First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal. Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomitant changes in the values of $Z_K$. Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data. We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect. \hfill\break

Based on the effect size $Z_K$, we then proposed a two-sample test, which provides a statistical means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another. Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $Kappa$ [e.g.,  @Blomberg_et_al2003]. However, $Kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information]. Thus, $Kappa$ should not be considered a standardized effect size that measures the strength of phylogenetic signal on a common scale. By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets. Thus when viewed from this perspective, the approach developed here aligns well with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons. As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.  \hfill\break

One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012]. Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation. This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$. For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$). In fact, our empirical example provided such a comparison, as SA:V is represented in $mm^{-1}$ while relative body size is a unitless ratio ($\frac{BW}{SVL}$). Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a]. Furthermore, tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal among datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence and observed variation are taken into account in the generation of the empirical sampling distribution via permutation.    \hfill\break

**Finally... Need Closing paragraph.**

### 2020-05-06 Manuscript/SupportingInfo.Rmd
Supp INfo

---
# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate

**FOR DCA**

1: balanced/symmetric trees
2: Kappa and K_Z
3: Beta from PGLS

@@

*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

@@

\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break

## Literature Survey

\newpage

# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>

>>>>

---
# Supporting Information for: A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate

@@

*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied most dramatically in simulations with fewer species and at intermediate values of lambda (Fig. S1, S2). Pectinate trees showed an interesting tendency to underestimate $\lambda$ across input values. Most dramatically, some simulations on the largest pectinate phylogeny ($n = 1024$) estimated $\lambda = 0$ for input values as high as 0.5 (Fig. S2).

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS1.png")

@@

\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees.

## Simulations of phylogenetic regression and ANOVA

We analyzed the precision of parameter estimation when parameters were obtained in phylogenetic regression and ANOVA. This involved simulations of dependent variables with $\lambda$ values ranging from 0 to 1 (21 intervals of 0.05 units) across pure-birth trees ($n=2^5 - 2^{10}$). For each $\lambda$ input value, independent variables were then generated with a known relationship to the simulated dependent variable ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.

*Results* As with the results from the regression analyses in the main text, we found poor precision of $\lambda$ estimation in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig. S3). Slope was reliably estimated across tree sizes and known $\lambda$ values (Fig. S4).

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS3.png")

\textbf{Figure S3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic ANOVA. Results are from input $\beta$ values of $0.5$. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS4.png")

\textbf{Figure S4}. Precision of slope estimates across known input $\beta$ values of phylogenetic regression. Results shown include all input $\lambda$ values.

## Effect Sizes of Pagel's Lambda ($Z_\lambda$) and Kappa ($Z_K$)

We proposed the use of effect sizes (Z-scores) as a measure of the strength of phylogenetic signal. These were obtained from summary parameters of phylogenetic signal: $\lambda$ and $Kappa$. The precision of effect sizes $Z_\lambda$ and $Z_K$ are summarized in the main text, primarily for simulations where $n=32$. Here we present results from additional simulations.

*Results* Effect sizes for Pagel's $\lambda$ ($Z_\lambda$) scale nonlinearly with input levels of phylogenetic signal, and this pattern is consistent across all tree sizes examined (Figure S5). Further, the precision of $Z_\lambda$ varied widely across input levels of phylogenetic signal, with decidely greater variation (less precision) at stronger input levels of phylogenetic signal (Figure S5). By contrast, effect sizes for *Kappa* ($Z_K$) displayed a linear increase with increasing input levels of phylogenetic signal, which remained consistent across all tree sizes examined (Figure S6). Additionally, the precision of $Z_K$ remained constant across input levels of phylogenetic signal (Figure S6). Together these results imply that $Z_K$ is a more reliable estimate of the strength of phylogenetic signal.

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS5.png")

\textbf{Figure S5}. Variation in effect size (Z score) of Pagel's $\lambda$ across input $\lambda$ values on various tree sizes.  \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS6.png")

\textbf{Figure S6}. Variation in effect size (Z score) of $K$ across input $\lambda$ values on various tree sizes. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS7.png")

\textbf{Figure S7}. Estimates of phylogenetic signal using $Kappa$ across known lambda inputs on various tree sizes.  Plot shows that while $Kappa$ increases with increasing phylogenetic signal, the precision of this estimate also varies. Thus, conversion to a standardized effect size ($Z_K$) is preferred.
